---
title: "Case_Study_1"
author: "The Miner League"
date: "June 19, 2019"
output: 
  html_document:
    keep_md: TRUE
---

#Introduction
The intent of the following outputs is to provide information related to analysis of beers and brewing within the United States. The team has investigated various items based on data sets that have been provided. The team has broken the analysis out into specific questions that needed to be answered. 

##Additional Notes
- All necessary files to recreate this analysis are located in the following GitHub repository: https://github.com/jbcroom/DS6306-MinerLeague-CaseStudies
- The markdown contains the ability for the user to set their working path to ensure data can be executed in a reproducible fashion. 
- The markdown requires the user have the dplyr and ggplot2 libraries installed.

###Initial Code Setup - The code chunk below is for setup of the environment and loading in the required data files
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r message=FALSE, warning=FALSE}
#commented out code below provides the necessary code to install required packages is needed
#install.packages("dplyr")
#install.packages("ggplot2")

#load necessary libraries
library(dplyr)
library(ggplot2)

#Setup the environment for working. Change the WorkingDirPath variable as needed to set the appropriate working directory
#WorkingDirPath = "C:/Users/sojha.WVOFFICE.000/Documents/SMU/Summer 2019/Doing Data Science/CaseStudy1_2_2_2"
WorkingDirPath = "C:\\Users\\croomb\\OneDrive - BAT\\Desktop\\Personal Training\\SMU\\DS 6306 - Doing Data Science\\DS6306-MinerLeague-CaseStudies\\CaseStudy1"
setwd(WorkingDirPath)

#Set variables for the file locations. Assumption here that the existing GitHub structure will be used.
beerFile = ".\\Data\\Beers.csv"
breweryFile = ".\\Data\\Breweries.csv"

#Read the files into dataframes
beerDF = read.csv(beerFile)
breweryDF = read.csv(breweryFile)
```

***

####Question 1 : How many breweries are present in each state?
```{r, fig.width=10, fig.height=10,message=FALSE, warning=FALSE}
#Counting the number of breweries by state. Requrn them in descending count order
brewerysByState = breweryDF %>% count(State,sort=TRUE)

#Add human redable names to the brewerysByState dataframe
names(brewerysByState) = c("State", "Count")

#Print out the listing of breweries and counts. NOTE: since DC is in the list there are 51 total entries
print(brewerysByState,n=51)

#Ploting the number of breweries in each state to provide a visual output
ggplot(data=brewerysByState, aes(x=reorder(State,Count), y=Count, fill=State)) + geom_bar(stat="identity", show.legend = FALSE) + labs(x ="State",y="Brewery Count",title="Number of Breweries by State") + theme(plot.title = element_text(size = 16,face= "bold",hjust = 0.5)) + theme(axis.text = element_text(size=8),axis.title=element_text(size=10))
```

#### As shown in the table above there are breweries located in all 50 states and Washington DC. Colorado (CO) has the most breweries with 47, while North Dakota (ND), South Dakota (SD), Washington, DC (DC) and West Virginia (WV) have the least number of breweries with 1. The graph provides an additional visualization of the data. 

***

####Question 2 : Merge beer data with the breweries data.  Print the first 6 observations and the last six observations to check the merged file.
```{r message=FALSE, warning=FALSE}
#Merging the beer and breweries data into a new data frame
BeersAndBreweriesDF = merge(x=beerDF,y=breweryDF,by.x="Brewery_id",by.y="Brew_ID")

#Cleanup the names of the columns to make the data more understandable
names(BeersAndBreweriesDF) = c("Brewery_id","Beer_Name","Beer_ID","ABV","IBU","Beer_Style","Beer_OZ","Brewery_Name","Brewery_City","Brewery_State")

#Display the first 6 and last 6 entries to confirm merge was successful
head(BeersAndBreweriesDF,6)
tail(BeersAndBreweriesDF,6)
```
#### As shown in the tables above we can confirm that the data has been merged successfully via looking at the first and last 6 records of the data set.

***

####Question 3 : Report the number of NA's in each column.
```{r}
#Sum the NAs in the datafram and display the total counts for all columnt
sapply(BeersAndBreweriesDF, function(x) sum(is.na(x)))
```
#### The table above shows the number of NAs (or missing values) in the data. In this instance there are 62 missing values in the ABV column and 1005 missing values in the IBU column. 

***


####Question 4 : Compute the median alcohol content and international bitterness unit for each state.  Plot a bar chart to compare.
```{r, fig.width=10, fig.height=10,message=FALSE, warning=FALSE}
#Calculating the median ABV and median IBV by state. Ignoring missing values
ABVByState = BeersAndBreweriesDF %>% group_by(Brewery_State) %>% summarise(median=median(ABV,na.rm=TRUE))
IBUByState = BeersAndBreweriesDF %>% group_by(Brewery_State) %>% summarise(median=median(IBU,na.rm=TRUE))

#Cleanup the dataframe names
names(ABVByState) = c("State","MedianABV")
names(IBUByState) = c("State","MedianIBU")

#Remove any extraneous NAs from the data
ABVByState = na.omit(ABVByState)
IBUByState = na.omit(IBUByState)

#Bar charts for comparison
#ABVByState
 ggplot(data=ABVByState, aes(x=reorder(State,MedianABV), y=MedianABV, fill=State)) + geom_bar(stat="identity", show.legend = FALSE) + labs(x ="State",y="ABV",title="ABV by State") + theme(plot.title = element_text(size = 16,face= "bold",hjust = 0.5)) + theme(axis.text = element_text(size=8),axis.title=element_text(size=10))
 
#IBUByState
  ggplot(data=IBUByState, aes(x=reorder(State,MedianIBU), y=MedianIBU, fill=State)) + geom_bar(stat="identity", show.legend = FALSE) + labs(x ="State",y="IBU",title="IBU by State") + theme(plot.title = element_text(size = 16,face= "bold",hjust = 0.5)) + theme(axis.text = element_text(size=8),axis.title=element_text(size=10))
```

#### The two graphs, ABV by State and IBU by State, depict the median alcohol content (ABV) and international bitterness unit (IBU) values for each state. It can be seen from the ABV by State graph that Kentucky (KY) has the highest median ABV while Utah (UT) has the lowest ABV. In reviewing the IBU graph, it can be seen that Maine (ME) has the highest median IBU value while Wisconsin (WI) has the lowest median IBU value.  

***

####Question 5 : Which state has the maximum alcoholic(ABV) beer? Which state has the most bitter (IBU) beer ?

```{r, fig.width=10, fig.height=10}
 #Filter the data to get the single row with maximum ABV
 maxABVStateDF = BeersAndBreweriesDF %>% filter(ABV == max(ABV,na.rm=TRUE))
 
 #Put just the Brewery State in a vector
 maxABVState = maxABVStateDF$Brewery_State[1]
 as.character(maxABVState)
 
 
 #Filter the data to get the single row with maximum IBU
 maxIBUStateDF = BeersAndBreweriesDF %>% filter(IBU ==   max(IBU,na.rm=TRUE))
 
 #Put just the Brewery State in a vector
 maxIBUState = maxIBUStateDF$Brewery_State[1]
 as.character(maxIBUState)
```
#### Based on the information provided above Colorado (CO) has the highest ABV of all states in the data set. Oregon (OR) has the most bitter beer (highest IBU) of all states in the data set.  

***


####Question 6 : Summary Statistics for the ABV variable.
```{r}
 #summary statistics for the ABV variable
 summary(BeersAndBreweriesDF$ABV)
```
#### Based on the table above we can see that the mean ABV is 0.05977 with a minimum of 0.00100 and a maximum of 0.12800. There are 62 NAs in the data as well. 
  
***

####Question 7 : Is there an apparent relationship between the bitterness of the beer and its alcholic content? Draw a scatter plot.

```{r, fig.width=10, fig.height=10}
#Scatter plot of bitterness (IBU) and alcoholic content (ABV)
relationship = ggplot(data = BeersAndBreweriesDF,
                      aes(x = ABV, y=IBU)) +
                      geom_point() +
                      geom_smooth(method="lm", color="red", linetype=2) +
                      labs(title="Alcoholic Content vs Bitterness",
                               x="Alcohol by Volume", y="International Bitterness Units")
relationship

#check correlation:`p-value < 2.2e-16` means 'REJECT the null hypothesis' that true correlation is equal to 0. We can conclude that alcoholic content (ABV) and bitterness (IBU) are significantly correlated with a correlation coefficient of `0.67` and p-value of `2.2*10^{-16}`.

cor.test(BeersAndBreweriesDF$ABV,BeersAndBreweriesDF$IBU)
```
#### Based on the data above it can be determined that there is a relationship between the bitterness of the beer and its alcholic content. It can concludes that alcoholic content (ABV) and bitterness (IBU) are significantly correlated with a correlation coefficient of `0.67` and p-value of `2.2*10^{-16).

